---
title: "PoCS Assignment 1"
author: "P. Alexander Burnham"
date: "September 4, 2018"
output: pdf_document
---
PoCS Assignment github repo: https://github.com/alexburn17/BurnhamPoCS




Worked with: Andrew, Alice and Chris

**#1)**

*Assumptions (including the information from assumptions a, b and f):*

$D_f \propto V^2l^2$  (assumption c)

$P \propto D_fV$  (assumption d)

$N \propto l^3$  (assumption e)

$P \propto N$  (assumption g)

*Operation:*

1) Substitute $D_f \propto V^2l^2$ into $P \propto D_fV$ then substitute $P$ for $l^3$ using assumptions g and e:

$$l^3 \propto V^3l^2$$  

2) Divide $l^3$ by $l^2$, rearrange assumption e ($l \propto N^{1/3}$) and substitute in after division: 

 $$l^3/l^2 \propto (V^3l^2)/l^2$$ 

$$N^{1/3} \propto V^3$$

3) Take the cube root of both sides and simplify: 

 $$\sqrt[3]{N^{1/3}} \propto \sqrt[3]{(V^3}$$  

 $$V \propto N^{1/9}$$ 

**#2)**
1/9 roughly equates to a slope of 0.111. After importing the current times for the 2000m (http://www.worldrowing.com/events/statistics/) and fitting a linear model as done in the McMahon and Bonner paper, we see a similar trend with a slope of 0.166. This is a slightly steeper slope but based on the coarseness of the data and the limited range of the variables in terms of orders of magnitude, they can be considered to be very similar. This slight increase, if more than a product of the limited data, might be due to technological advancements that allow for the reduction of friction in larger hulled boats. A premise (at least of sailing) is that larger boats go faster all else being equal so technological advancement might have strengthened this relationship.


```{r, echo=FALSE, message=FALSE}
###################################################################
# PoCS Assignment 1 code
# P. Alexander Burnham
# Sept 5, 2018
###################################################################

#Preliminaries:
# Clear memory of characters

rm(list=ls())


# set working Directory:
setwd("~/Documents/GitHub/BurnhamPoCS")

# Read in Data:
rowing <- read.csv("assignment1_rowing.csv", 
                header=TRUE, 
                sep = ",", 
                stringsAsFactors = FALSE)

# convert time into seconds
rowing$seconds <- as.numeric(strptime(rowing$Time, format="%M:%S") - as.POSIXct(format(Sys.Date())), units="secs")

# convert to meters per second
rowing$MperSec <- 2000/rowing$seconds

# model fitting line to data set
mod <- lm(rowing$MperSec~rowing$Class)

# plot speed by number of oarsman 
plot(rowing$Class, rowing$MperSec, xlab="Number of Oarsman", ylab="speed (m/s)", main = "Modern World Records for the 2000m by Class")
abline(mod, col = "red")

mod$coefficients # print model


```


**#3)**

a) The scale does seem to match the 2/3 scaling hypothesis. The data were taken from (https://www.iwf.net/results/world-records/) body mass and weight lifted were log base 10 transformed and the slopes of the line of best fit for each regression were close to 0.666 (Clean and Jerk = 0.609, Snatch = 0.601 and Total = 0.597). 



```{r, echo=FALSE}
###################################################################

# Read in Data:
weights <- read.csv("assigment1_weights.csv", 
                header=TRUE, 
                sep = ",", 
                stringsAsFactors = FALSE)

weights$logMass <- log10(weights$BodyMass)
weights$logWeight <- log10(weights$Weight)
# split by three classes
weightssplit <- split(weights, weights$Type)

par(mfrow=c(1,3))

# model for clean and jerk
mod1 <- lm(weightssplit$cleanJerk$logWeight ~ weightssplit$cleanJerk$logMass)

# plot clean and jerk 
plot(weightssplit$cleanJerk$logMass, weightssplit$cleanJerk$logWeight, xlab="log(Weight Class (kg))", ylab="log(Weight Lifted (kg))", main = "Record for Clean & Jerk")
abline(mod1, col = "red")


# model for snatch 
mod2 <- lm(weightssplit$snatch$logWeight ~ weightssplit$snatch$logMass)

# plot Snatch
plot(weightssplit$snatch$logMass, weightssplit$snatch$logWeight, xlab="log(Weight Class (kg))", ylab="log(Weight Lifted (kg))", main = "Record for Snatch")
abline(mod2, col = "blue")


# model for total
mod3 <- lm(weightssplit$Total$logWeight ~ weightssplit$Total$logMass)

# plot total
plot(weightssplit$Total$logMass, weightssplit$Total$logWeight, xlab="log(Weight Class (kg))", ylab="log(Weight Lifted (kg))", main = "Record for Total")
abline(mod3, col = "green")

par(mfrow=c(1,1))


```

Intercepts and Coefficients for Clean and Jerk, Snatch and Total Respectivly 
```{r, echo=FALSE}
mod1$coefficients
mod2$coefficients
mod3$coefficients

```


b) Using the residuals from the Total catagory, in order to determine overall scaled world champion, we see that the lift fomr the champion in the highest weight class still holds (TALAKHADZE Lasha with 477kg). However, the second place scaled champion is in the 3rd lowest weight class (LIAO Hui with 359 kg).


```{r, echo=FALSE}


# plot residuals to determine who punches above their weight
plot(weightssplit$Total$logWeight, mod3$residuals, ylab = "Residuals", xlab = "log(Weight Class)", main = "Total World Record Rescaled")
abline(h=0, col="red")

```

**#4)**

Find the Dimensions:

$q_1: [l] = L$

$q_2: [m] = M$

$q_3: [g] = LT^{-2}$

$q_4: [\tau] = T$

Set up dimensions to find $\pi_1$

$\pi_1 = (L)^{x_1}(M)^{x_2}(LT^{-2})^{x_3}(T)^{x_4}$

$L: x_1+x_3=0$

$M: x_2=0$

$T: -2x_3+x_4=0$

One dimensionless parameter (4-3=1). Equation is in one space so make assumptions that $x_1=1$.

Solve algerbraicially after substitution:

$L: 1+x_3=0$ thus $x_3=-1$

$M: x_2=0$

$T: -2x_3+x_4=0$ substitute $-1$ in for $x_3$ to get $x_4=-2$

$$x_1=1,   x_2=0,   x_3=-1,   x_4=-2$$

$$\pi_1 = l^1m^0g^{-1}\tau^{-2}$$

$$\pi_1 = l/(g\tau^{2})$$ 

multiply both sides by $g$, divide by $l$ and take the square root of both sides

$$\sqrt{\tau^2} \propto \sqrt{l/g}$$

$$\tau \propto \sqrt{l/g}$$

**#5)**

Find the dimensions:

$q_1: [b] = L^2/(T^3)$

$q_2: [\rho] = M/L^3$

$q_3: [V] = L/T$

$q_4: [l] = L$

$q_5: [\sigma] = M/LT^2$


Set up dimensions to find $\pi_1$:

$\pi_1 = (L^2/(T^3))^{x_1}(M/L^3)^{x_2}(L/T)^{x_3}(L)^{x_4}(M/LT^2)^{x_5}$

$L: 2x_1-3x_2+x_3+x_4-x_5=0$

$M: x_2+x_5=0$

$T: -3x_1-x_3-2x_5=0$

Two dimensionless parameters (5-3=2). Equation is in two space so make assumptions that $x_1=1, x_2=1$.


Solve for $x_i$ algebraically:

$M: (1)+x_5=0$ thus $x_5=-1$ and substitute into equation for T

$T: -3(1)-x_3-2(-1)=0$ thus $x_3 = 1$ and substitue into equation for L

$L: 2(1)-3(1)+(1)+x_4-(-1)=0$ thus $x_4 = -1$ 

Solutions:

$x_1=1,   x_2=1,   x_3=1,   x_4=-1,   x_5=-1$

Substitute in parameters and exponents:

$$\pi_1 = b^1\rho^1V^1l^{-1}\sigma^{-1}$$

$$\pi_1 = Vb\rho/(l\sigma)$$

multiply both sides by $\sigma/(b\rho)$

$$V/l \propto \sigma/(b\rho)$$



**#6)**
Find the Dimensions:

$q_1: [E] = ML^2/(T^2)$

$q_2: [P] = M/L^3$

$q_3: [R] = L$

$q_4: [t] = T$

Set up dimensions to find $\pi_1$

$\pi_1 = (ML^2/(T^2))^{x_1}(M/L^3)^{x_2}(r)^{x_3}(T)^{x_4}$

$L: 2x_1-3x_2+x_3=0$

$M: x_1+x_2=0$

$T: -2x_1+x_4=0$

One dimensionless parameter (4-3=1). Equation is in one space so make assumptions that $x_1=1$.

Solve algerbraicially after substitution:

$L: 2-3x_2+x_3=0$

$M: 1+x_2=0$ thus $x_2 = -1$

$T: -2+x_4=0$ thus $x_4 = 2$

plug in $x_2$ into the equation for L in order to get all exponants:

$$x_1=1,   x_2=-1,   x_3=-5,   x_4=2$$

$$\pi_1 = E^1P^{-1}R^{-5}t^2$$

Multiply to get E to one side then Flip denominators 

$$\pi_1 = E^1t^2/(P^1R^5)$$

$$E = pR^5/t^2$$ 


**#7)**

a)

Find the dimensions:

$q_1: [G] = L^3/(MT^2)$

$q_2: [M] = M$

$q_3: [m] = M$

$q_4: [\tau] = T$

$q_5: [r] = L$

b)

Set up dimensions to find $\pi_1$:

$\pi_1 = (L^3/(MT^2))^{x_1}(M)^{x_2}(m)^{x_3}(T)^{x_4}(L)^{x_5}$

$L: 3x_1+x_5=0$

$M: -x_1+x_2+x_3=0$

$T: -2x_1+x_4=0$

Two dimensionless parameters (5-3=2). Equation is in two space so make assumptions that $x_1=1, x_2=1$.

c)

Solve for $x_i$ algebraically to get:

$x_1=1,   x_2=1,   x_3=0,   x_4=2,   x_5=-3$

Substitute in parameters and exponents:

$$\pi_1 = G^1M^1m^0\tau^2r^{-3}$$

Change sides and remove constants

$$\pi_1 = GM\tau^2/(r^{-3})$$

$$r^3 \propto \tau^2$$

d)

Not supprisingly, Kepler's 3rd law holds up well. Plotting the period squared against the radius cubed creates a near perfect regression (R^2^=1 and slope = 0.999). This worked for 8 planets in our solar system as well as one of the thousands of small frozen non-planets. This indicates that Kepler's work in the orbital mechanics of planets generalizes to other things in orbits.

```{r, echo=FALSE}
# set up data
planet <- c("Mercury", "Venus", "Earth", "Mars", "Jupiter", "Saturn", "Uranus", "Neptune", "Pluto")
r <- c(0.387, 0.723, 1, 1.524, 5.203, 9.537, 19.19, 30.07, 39.48)
tau <- c(0.241, 0.615, 1, 1.881, 11.86, 29.42, 83.75, 163.7, 248.0)
r3 <- r^3
tau2 <- tau^2

# plot and model data
plot(log10(r3), log10(tau2), col="white", ylab = "log(Period (T^2))", xlab = "log(Radius (r^3))", main = "Test of Kepler's 3rd Law")
mod5 <- lm(log10(tau2)~log10(r3))
text(x=log10(r3), y= log10(tau2), labels = planet)
abline(mod5, col="red")

mod5$coefficients
```

**#8)**

a)

Show equations in matrix notation:

$$\begin{bmatrix} L_1\\ L_2\\ L_3\\    \end{bmatrix} = \begin{bmatrix} {c_1}^{-1}\\ {c_2}^{-1}\\ {c_3}^{-1}\\    \end{bmatrix}  \begin{bmatrix} V^{\gamma_1}\\ V^{\gamma_2}\\ V^{\gamma_3}\\    \end{bmatrix}$$

Substitute:

$$L_1L_2L_3 = (c_1c_2c_3)^{-1}V^{\gamma_1+\gamma_2+\gamma_3}$$

Contstants and exponants are equal to 1 by definition:

$$L_1L_2L_3 = (1)V^{1}$$

$$L_1L_2L_3 = V$$

b)

For an isometric scaling $\gamma_i=1/3$ so all of the 3 dimensions grow at the same rate

c)

$$S = 2(L_1L_2)+ 2(L_2L_3)+ 2(L_1L_3)$$

d)

Assume, as directed, isometric $\gamma_i$ and set constants equal to one to see general relationship. We see that as volume increases, surface area increases at a decaying rate and the surface area to volume ration decreases as a negative exponential function, Making volume a dominate variable in this function.

```{r, echo=FALSE}

Volume <- 1:1000

SurfVol <- function(V=Volume, 
                    g1 = 1/3, g2 = 1/3, g3 = 1/3,
                    c1 = 1, c2 = 1, c3 = 1){
# equations:
L1 <- c1 * V^g1
L2 <- c2 * V^g2
L3 <- c3 * V^g3

S <- 2*(L1*L2) + 2*(L1*L3) + 2*(L3*L3)

mat <- cbind(V, L1, L2, L3, S)
mat <- data.frame(mat)

return(mat)

}

mat <- SurfVol()

par(mfrow=c(1,2))

plot(mat$V, mat$S, ylab = "S", xlab="V", main = "Surface Area by Volume")

plot(mat$V, mat$S/mat$V, ylab = "S/V", xlab="V", main = "Surface Area to Volume Ratio")

par(mfrow=c(1,1))

```


e)
The fastest set is when $\gamma_3=1$ and the other two are 0 as is permissable by $\gamma_1+\gamma_2+\gamma_3=1$ & $\gamma_1\leq\gamma_2\leq\gamma_3$ and is slowest when the smallest allowable $\gamma_3$ value is inputted, $1/3$. 
```{r, echo=FALSE}
par(mfrow=c(1,2))

mat <- SurfVol(g1 = 0, g2 = 0, g3 = 1)
plot(mat$V, mat$S, ylab = "S", xlab="V", main = "Surface Area by Volume MAX")

mat <- SurfVol(g1 = 1/3, g2 = 1/3, g3 = 1/3)
plot(mat$V, mat$S, ylab = "S", xlab="V", main = "Surface Area by Volume MIN")

par(mfrow=c(1,1))
```

