---
title: "PoCS Assignment 5"
author: "P. Alexander Burnham"
date: "October 2, 2018"
output: pdf_document
---
PoCS Assignment github repo: https://github.com/alexburn17/BurnhamPoCS


Worked with: Yu Han, Edison & Kewang


**Problem 1:**

**a)**

Step 1: sub in $1/D$:

$$S=\sum_{i=1}^{n}p^2_i$$

$$\sum_{i=1}^{n}p^2_i = \sum_{i=1}^{D}(\frac{1}{D})^2_i$$

Step 2: Do summation and simplify:

$$S=D(\frac{1}{D})^2$$

$$S=\frac{1}{D}$$

$$D=\frac{1}{S}$$

Step 3: Sub in $S$:

$$D=\frac{1}{\sum_{i=1}^{n}p^2_i}$$

**b)**

Step 1: sub in  from part a $(S=1/D)$

$$G\equiv 1-S = 1-\sum_{i=1}^{n}p^2_i$$

$$G\ = 1-\frac{1}{D}$$

$$D = \frac{1}{1-G}$$

Step 2: sub in definition of G

$$D = \frac{1}{1-(1-S)}$$

$$D = \frac{1}{S}$$

Step 3: sub in S

$$D = \frac{1}{\sum_{i=1}^{n}p^2_i}$$

In terms of $D$, these two metrics are identical. 

**c)**

Part 1: sub in $1/D$

$$H=-\sum_{i=1}^{n}p_iln(p_i)$$

$$H=-\sum_{i=1}^{D}(\frac{1}{D})ln(\frac{1}{D})$$

Part 2: cancel

$$H=-D(\frac{1}{D})ln(\frac{1}{D})$$

$$H=ln(D)$$

$$D=e^H$$

Step 3: sub $H$ back in

 $$D=e^{-\sum_{i=1}^{n}p_iln(p_i)}$$

**d)**

Step 1: sub in $1/D$

$$H_q^{(R)}=\frac{1}{q-1}(-ln\sum_{i=1}^{n}p_i^q)$$

$$H_q^{(R)}=\frac{1}{q-1}(-ln\sum_{i=1}^{D}(\frac{1}{D})^q)$$

$$H_q^{(R)}=\frac{1}{q-1}(-lnD(\frac{1}{D})^q)$$

Step 2: simplify 

$$H_q^{(R)}=\frac{1}{q-1}(-lnD^{(1-q)})$$

$$H_q^{(R)}=\frac{1}{q-1}(lnD^{(q-1)})$$

Step 3: Multiply both sides by $q-1$ and simplify

$$H_q^{(R)}(q-1)=lnD(q-1)$$

$$H_q^{(R)}=lnD$$

$$D=e^{H_q^{(R)}}$$

Step 4: plug in $H_q^{(R)}$

$$D=e^{\frac{1}{q-1}(-ln\sum_{i=1}^{n}p_i^q)}$$



**e)**

Step 1: plug in $1/D$ and simplify 

$$H_q^{(T)}=\frac{1}{q-1}(1-\sum_{i=1}^{n}p_i^q)$$

$$H_q^{(T)}=\frac{1}{q-1}(1-\sum_{i=1}^{D}(\frac{1}{D})^q)$$

$$H_q^{(T)}=\frac{1}{q-1}(1-D(\frac{1}{D})^q)$$

$$H_q^{(T)}=\frac{1}{q-1}(1-D^{(1-q)})$$

Step 2: rearrange and simplify 

$$D=((1-(q-1))H_q^{(T)})^{\frac{1}{{(1-q)}}}$$

Step 3: sub in $H_q^{(T)}$

$$D=((2-q)(\frac{1}{q-1}(1-\sum_{i=1}^{n}p_i^q)))^{\frac{1}{{(1-q)}}}$$

Both of these depend on Shannon's E

**f)**



$$\lim_{n\to\infty} (1+\frac{H_q^{(T)}}{n})^n=e^{H_q^{(T)}}$$

$$\lim_{q\to1} (1+\frac{H_q^{(T)}}{\frac{1}{1-q}})^{\frac{1}{1-q}}=e^{H_q^{(T)}}$$

$$D=((1-(q-1))H_q^{(T)})^{\frac{1}{{(1-q)}}}$$

$$D=1+(\frac{H_q^{(T)}}{\frac{1}{1-q}})^{\frac{1}{1-q}}$$

$$\lim_{q\to1} = 1+(\frac{H_q^{(T)}}{\frac{1}{1-q}})^{\frac{1}{1-q}}=D$$

$$D=e^{H_q^{(T)}}$$

This matches shannon's entropy $D=e^H$



**Problem 2:**
Step 1: Set up equation and sub in

$$\Psi=F+\lambda G$$

$$\Psi=\frac{C}{H}+\lambda G$$

$$=\frac{\sum_{i=1}^{n}p_i ln(j+a)}{-g\sum_{i=1}^{n}p_i ln p_i}+\lambda\sum_{i=1}^{n}p_i-1$$

Step 2: take partial derivative 

$$\frac{\partial \Psi}{\partial p_j}=\frac{ln(j+a)H + Cg(Hlnp_i))}{H^2}+\lambda=0$$

$$\frac{\partial \Psi}{\partial p_j}=\frac{ln(j+a)(-g\sum_{i=1}^{n}p_i ln p_i)-(\sum_{i=1}^{n}p_ih(j+a)(-g(Hlnp_i))}{(-g\sum_{i=1}^{n}p_i ln p_i)^2}+\lambda$$

Step 3: simplify and rearrange

$$-\lambda H^2=Hln(j+a)+Cg(Hlnp_i)$$

$$p_i=e^{\frac{-\lambda H^2-Hln(j+a)-Cg}{Cg}}$$

$$p_i=e^{-\lambda H^2/Cg}(j+a)^{-H/Cg}$$

Step 4: Find expression connecting all variables while also showing that $e^{-\lambda H^2/Cg}=1$

$$p_i=e^{-1-\frac{\lambda H^2}{Cg}(a+j)^{\frac{-H}{Cg}}}$$

take the natural log:

$$lnp_i={-1-\frac{\lambda H^2}{Cg}+{\frac{-H}{Cg}}ln(a+j)}$$

$$H=-g\sum_{i=1}^{n}p_iln(p_i)$$

$$H=-g\sum_{i=1}^{n}p_i\Bigg[\Big(-1-\frac{\lambda H^2}{Cg}\Big)+\Big(\frac{-H}{Cg}\Big)ln(a+j)  \Bigg]$$

$$H=g\Bigg[\sum_{i=1}^{n}p_i\Big(\frac{1+\lambda H^2}{Cg}\Big)+\frac{H}{C}\sum_{i=1}^{n}p_iln(a+j)  \Bigg]$$

$\sum_{i=1}^{n}p_i$ goes to 1 and $\sum_{i=1}^{n}p_iln(a+j)$ is C

$$H=g+\frac{\lambda H^2}{C}+H$$

$$0=g+\frac{\lambda H^2}{C}$$

$$-g=\frac{\lambda H^2}{C}$$

Find expression connecting all variables….

$$\lambda=\frac{-gC}{H^2}$$

Step 5: plug $\lambda=\frac{-gC}{H^2}$ into $p_i=e^{-1-\frac{\lambda H^2}{Cg}(a+j)^{\frac{-H}{Cg}}}$

$$p_i=e^0(a+j)^{\frac{-H}{Cg}}$$

$$p_i=(a+j)^{\frac{-H}{Cg}}$$
**Problem 3:**

**A)**

Use this expression for large N for various alphas between 1 and 2. Iterate through the summation:
$$\sum_{j=1}^{N}(j+1)^{-\alpha}-1=0$$



```{r, echo=FALSE, message=FALSE}
# Problem 3

# a)

# create j
N <- 1000


# sequnce of alpha
alpha <- seq(1,2,length.out=1000)
fa <- matrix(nrow=N, ncol=length(alpha))

# function of a
for(a in 1:length(alpha)){
  for(j in 1:N){
  
    fa[j, a] <- ((j+1)^(-alpha[a]))

}
}

y <- colSums(fa)-1

plot(x=alpha, y=y,  ylab = "function(alpha)", xlab="alpha", main = "alpha by function of alpha to determine alpha")
abline(h=0, col = "red")
text(x=1.8, y=5, "alpha = 1.723724")

df <- data.frame(alpha, y)

```


**B)**

$$\sum_{i=1}^{n}(j+a)^{-1}=1$$

$$\sum_{i=1}^{n}\frac{1}{i+a}=1$$

approximation:

$$\int_{1}^{n} \frac{1}{x+a}dx=1$$ 

antiderivative: 

$$ln(a+x)|^{n}_{1}=1$$

$$ln(a+n)-ln(a+1)=1$$

$$\frac{a+n}{a+1}=e$$

$$a=\frac{e-n}{1-e}$$

as n goes to infinity, a goes to positive infinity 



**Problem 4 (A, B):**

```{r, echo=FALSE, message=FALSE}
setwd("~/Documents/GitHub/BurnhamPoCS/assignment3")


# Read in Data:
wordsFreq <- read.table("google_vocab_rawwordfreqs.txt", 
                        header=TRUE, 
                        sep = " ", 
                        stringsAsFactors = FALSE)



# Problem 4

# a)
k <- c(1:199)
N_k <- 3.46 * (10^8)*k^(-0.661)
plot(y=log10(N_k), log10(k), ylab="log10(N_k)", xlab = "log10(k)", main = "Extrapolated google word plot (k=1-199)")

# b)

fullset <- c(N_k, wordsFreq$X95119665584)

paste("The mean is:", mean(fullset, na.rm=T)) # mean
paste("The variance is:", (sd(fullset))^2) # varaince


```


**C)**

$$i =\frac{N_1}{\sum k * N_k (total)}$$

$$ii =\frac{\sum N_k (extrapolated)}{\sum N_k (total)}$$

$$iii =\frac{\sum N_k * k (extrapolated)}{\sum k * N_k (total)}$$


```{r, echo=FALSE, message=FALSE}


# c)
# i sum(nk * k)
Denom <- sum(c(1:length(fullset))*fullset)

paste("The answer to C part i is:", N_k[1]/Denom)

# ii

paste("The answer to C part ii is:", sum(N_k)/sum(fullset))

# iii

paste("The answer to C part iii is:", sum(N_k*k)/Denom)

```


